Title : Reproducible Research project 1

STEP 1 : Loading and Reprocessing data

```{r}
setwd("C:/NelsonRepo/Technical/DataScience/R_Programming/Reproducible Research/Week2Project/repdata")
InputData <- read.csv('activity.csv')

```



STEP 2 :  Mean total number of steps taken per day


Ignore missing NA value

```{r}
InputDataNoNA <- InputData[complete.cases(InputData),]
```

    1. Number of steps taken each day
    
    ```{r}
    StepsDaySum <- aggregate(InputDataNoNA["steps"], by=InputDataNoNA["date"], FUN=sum)
    StepsDaySum
    ```
    
    2. Histogram for total number of steps taken each day
    
    ```{r}
    str(StepsDaySum)
    library(ggplot2)  
    ggplot(StepsDaySum, aes(x=steps)) + geom_histogram(fill = "blue", binwidth = 1000) + labs(title = "Daily Steps", x = "Total Steps", y = "Frequency")
    
    ```
    
    3. Mean and Median of total steps taken each day
    
    ```{r}
    
   StepsDaymean <-  mean(InputDataNoNA$steps, na.rm=TRUE)
   StepsDaymean
   
   StepsDaymedian <-  median(InputDataNoNA$steps, na.rm=TRUE)
   StepsDaymedian
    
    ```
    
    4. Average daily activity pattern
    
    ```{r}
   library(dplyr)
        Avgactpat <- InputData %>% group_by(interval)%>% filter(!is.na(steps))%>%summarise(avgsteps=mean(steps,na.rm=TRUE))
   
   
    ```
   
   5. Which interval has maximum number of steps
   
  ```{r}
  
 Avgactpat[which.max(Avgactpat$avgsteps),]
 
 ```
 

STEP 3 :  Imputing missing values


      1.  Total number for missing values in the dataset
      
      
      ```{r}
      
      sum(is.na(InputData$steps))
      
      ```
      
      2. Input missing values
      
      
      ```{r}
      
      tmpInputData <- InputData
      
      Nulv <- is.na(tmpInputData$steps)
      avginterval <- tapply(tmpInputData$steps,tmpInputData$interval,mean,na.rm=TRUE,simplify=TRUE)
      
      tmpInputData$steps[Nulv] <- avginterval[as.character(tmpInputData$interval[Nulv])]
      
      
      
      ```
      3. Histogram for total number of steps taken each day and mean, median vaue
      
      
      ```{r}
      
      
      names(tmpInputData)
      
      TotalSteps2 <- tmpInputData %>% group_by(date) %>% summarise(totalsteps = sum(steps, na.rm=TRUE))
      names(TotalSteps2)
      
      ggplot(TotalSteps2,aes(x= totalsteps)) + geom_histogram(fill = "blue", binwidth = 1000) +
      labs(title = "Daily Steps including Missing values", x = "Interval", y = "No. of Steps")
      
      ```
      
      mean & Median
      
      ```{r}
      
      Mean_Steps2 <- mean(TotalSteps2$totalsteps,na.rm=TRUE)
      Mean_Steps2
      
      Median_Steps2 <- median(TotalSteps2$totalsteps,na.rm=TRUE)
      Mean_Steps2
      ```
  
  STEP4  :Differences in activity patterns between weekdays and weekends
  
  
     
     1. Categorize into weekdays and weekend
     
      ```{r}
        tmpInputData2 <- tmpInputData
        
        tmpInputData2$date<- as.Date(tmpInputData2$date)  ##covert chr to date
        
        tmpInputData2<- tmpInputData2 %>% 
        mutate(weektype= ifelse(weekdays(tmpInputData2$date)=="Saturday" | weekdays(tmpInputData2$date)=="Sunday","Weekend","Weekday"))
        
        head(tmpInputData2)
        
      ```
      
           
      2.  Plot the graphy for weekend and weekday
      
      
      ```{r}
      
       InputDatawktyp <- tmpInputData2 %>% group_by(interval,weektype) %>% summarise(ttlsteps = mean(steps,na.rm=TRUE))
       
       head(InputDatawktyp)
       
      
        plot<- ggplot(InputDatawktyp, aes(x =interval , y=ttlsteps, color=weektype)) +
               geom_line() +
               labs(title = "Avg. Daily Steps by Weektype", x = "Interval", y = "No. of Steps") +
               facet_wrap(~weektype, ncol = 1, nrow=2)
               
        print(plot)
        
        ```
       
       
      
      
      
        
      
      
      
      
      
      
      
      
      
      
      
 
 
   
  
   
   
  
